Enhancing Forecast Reconciliation: A Study of Alternative Covariance Estimators

### 2025-09-22 literature/paper_draft.qmd
add tables to illstrate tourism structure

---
title: "Enhancing Forecast Reconciliation: A Study of Alternative Covariance Estimators"
author: 
  - Vincent Su
  - Shanika Wickramasuriya (supv.)
  - George Athanasopoulos (supv.)
format: 
  pdf:
    include-in-header:
      text: |
        <!-- \usepackage{amsmath} -->
        <!-- \usepackage{mathspec} -->
        \usepackage{algorithm}
        \usepackage{algpseudocode}
        \usepackage[colorlinks=true, linkcolor=blue, citecolor=blue, urlcolor=blue]{hyperref}
        \usepackage{array}
    mathspec: true
    number-sections: true
# crossref:
#   custom:
#     - kind: float
#       key: alg
#       reference-prefix: Algorithm
#       caption-prefix: Algorithm
#       # for PDF output only, tell LaTeX to use the `algorithm` env:
#       latex-env: algorithm
execute:
  cache: false
  echo: false
  warning: false
bibliography: D:/Github/Recon_Honours_Thesis/references.bib
# bibliographystyle: apa
csl: D:/Github/Recon_Honours_Thesis/apa.csl
editor_options: 
  chunk_output_type: console
---

@@

```{r library}
library(ggplot2)
library(patchwork)
library(ReconCov)
library(dplyr)
library(tidyr)
library(fable)
library(fabletools)
library(feasts)
library(tsibble)
library(lubridate)

path <- "D:/Github/Recon_Honours_Thesis"
knitr::opts_knit$set(root.dir = path)
# setwd(path)
```

@@

<!-- - A table/figure to illustrate the hierarchy -->

@@

```{r}
library(stringr)

desired_order <- c(
  "Australia", "Australia_by_Purpose",
  "State",     "State_by_Purpose",
  "Zone",      "Zone_by_Purpose",
  "Region",    "Region_by_Purpose"
)
```

@@

# % improv plot by level
tourism_MSE |>
  group_by(series, h) |>
  mutate(
    base_MSE = MSE[.model == "base"]
  ) |>
  ungroup() |>
  group_by(.model, level, h) |>
  summarise(
    MSE = mean(MSE),
    base_MSE = mean(base_MSE),
    pct_change = (MSE - base_MSE) / base_MSE * 100
  ) |>
  filter(.model != "mint_sample") %>%
  # Reorder the levels according to the desired order
  mutate(
    level = factor(
      level,
      levels = desired_order,
      labels = str_replace_all(desired_order, "_", " ")  # replace "_" with " "
    )
  ) %>% 
  ggplot(aes(x = h, y = pct_change, color = .model)) +
    geom_line() +
    labs(x = "Horizon", y = "% improvements") +
    facet_wrap(~ level, ncol=2, scales = "free_y") +
    theme_bw() +
    scale_x_continuous(limits = c(1, 12), expand = c(0, 0)) +
    theme(legend.position = "bottom", legend.title = element_blank())
```

@@

visnights_full %>% 
  filter(
    is_aggregated(Purpose) & is_aggregated(Zone) & is_aggregated(Region),
    !is_aggregated(State)
  ) %>% 
  model(
    STL(Nights)
  ) %>% 
  components() %>% 
  pivot_longer(
    cols = c(Nights, trend, season_year, remainder, season_adjust),
    names_to = "component",
    values_to = "value"
  ) %>%
  filter(
    component == c("Nights", "trend"),
    year(Month) >= 2000
  ) %>%
  as_tibble() %>%
  mutate(component = recode(component,                     # give them nicer names
                            Nights = "Observed Nights",
                            trend  = "STL Decomposed Trend")) %>% 
  ggplot(aes(Month, value, color = as.character(State))) +
    geom_line() +
    facet_wrap(~ component, ncol=1, scales = "free_y") +
    labs(
      x = "Month",
      y = "Visitor Nights (in thousands)",
      color = "State"
    ) +
    theme_bw()
```

@@

# % improv plot by level
tourism_MSE_ex2016 |>
  group_by(series, h) |>
  mutate(
    base_MSE = MSE[.model == "base"]
  ) |>
  ungroup() |>
  group_by(.model, level, h) |>
  summarise(
    MSE = mean(MSE),
    base_MSE = mean(base_MSE),
    pct_change = (MSE - base_MSE) / base_MSE * 100
  ) |>
  filter(.model != "mint_sample") %>%
  # Reorder the levels according to the desired order
  mutate(
    level = factor(
      level,
      levels = desired_order,
      labels = str_replace_all(desired_order, "_", " ")  # replace "_" with " "
    )
  ) %>%
  ggplot(aes(x = h, y = pct_change, color = .model)) +
    geom_line() +
    labs(x = "Horizon", y = "% improvements") +
    facet_wrap(~ level, ncol=2, scales = "free_y") +
    theme_bw() +
    scale_x_continuous(limits = c(1, 12), expand = c(0, 0)) +
    theme(legend.position = "bottom", legend.title = element_blank())
```

@@

    "Final writing, presenation, submission"
  ),
  Deliverable = c(
    "Initial literature review + paper draft",
    "ReconCov package: Main codes for estimators + simulation framework + cross-validation",
    "NOVELIST initial simulation results + research proposal + presentation",
    "More simulation results on NOVELIST",
    "Empirical results on forecasting Tourism data",
    "New simulation design, POET & PC-adjusted NOVELIST simulation results",
    "Other estimators simulation results and full-fledged covariance estimators assessment",
    "Presentation and thesis manuscript"
  )
) %>%
  kable(
    format = "latex",
    booktabs = TRUE,
    longtable = TRUE,
    caption = "Project Timeline and Key Milestones",
    escape = TRUE
  ) %>%
  column_spec(1, width = "1.9cm") %>%
  column_spec(2, width = "8cm") %>%
  column_spec(3, width = "4.8cm") %>%
  kable_styling(latex_options = "hold_position", font_size = 10)
```

@@

>>>>

---
title: "Enhancing Forecast Reconciliation: A Study of Alternative Covariance Estimators"
author: 
  - Vincent Su
  - Shanika Wickramasuriya (supv.)
  - George Athanasopoulos (supv.)
format: 
  pdf:
    include-in-header:
      text: |
        <!-- \usepackage{amsmath} -->
        <!-- \usepackage{mathspec} -->
        \usepackage{algorithm}
        \usepackage{algpseudocode}
        \usepackage[colorlinks=true, linkcolor=blue, citecolor=blue, urlcolor=blue]{hyperref}
        \usepackage{array}
    mathspec: true
    number-sections: true
# crossref:
#   custom:
#     - kind: float
#       key: alg
#       reference-prefix: Algorithm
#       caption-prefix: Algorithm
#       # for PDF output only, tell LaTeX to use the `algorithm` env:
#       latex-env: algorithm
execute:
  cache: true
  echo: false
  warning: false
bibliography: D:/Github/Recon_Honours_Thesis/references.bib
# bibliographystyle: apa
csl: D:/Github/Recon_Honours_Thesis/apa.csl
editor_options: 
  chunk_output_type: console
---

@@

```{r library}
library(ggplot2)
library(patchwork)
library(dplyr)
library(tidyr)
library(fable)
library(fabletools)
library(feasts)
library(tsibble)
library(lubridate)
library(knitr)
library(kableExtra)
library(stringr)

library(ReconCov)

path <- "D:/Github/Recon_Honours_Thesis/"
knitr::opts_knit$set(root.dir = path)
# setwd(path)
```

@@

```{r}
#| tbl-cap: "Hierarchical and grouped structure of Australian domestic tourism flows"
#| label: tbl-tourism-structure

num_series_geo <- c(1, 7, 27, 77)
table_structure <- tibble(
  `Geographical division` = c("Australia", "States", "Zones", "Regions"),
  `Number of series per geographical division` = num_series_geo,
  `Number of series per purpose` = 4*num_series_geo,
  `Total number of series` = 5*num_series_geo
)
# add total per row
table_structure <- rbind(
  table_structure,
  c("Total", colSums(table_structure[,-1]))
)

# width margin smaller (90%)
table_structure |> 
  knitr::kable(
    align = "lccc",
    booktabs = TRUE
  )
```

@@

```{r}
library(stringr)
desired_order <- c(
  "Australia", "Australia_by_Purpose",
  "State",     "State_by_Purpose",
  "Zone",      "Zone_by_Purpose",
  "Region",    "Region_by_Purpose"
)
```

@@

# % improv plot by level
tourism_MSE |>
  group_by(series, h) |>
  mutate(
    base_MSE = MSE[.model == "base"]
  ) |>
  ungroup() |>
  group_by(.model, level, h) |>
  summarise(
    MSE = mean(MSE),
    base_MSE = mean(base_MSE),
    pct_change = (MSE - base_MSE) / base_MSE * 100
  ) |>
  filter(.model != "mint_sample") |>
  # Reorder the levels according to the desired order
  mutate(
    level = factor(
      level,
      levels = desired_order,
      labels = str_replace_all(desired_order, "_", " ")  # replace "_" with " "
    )
  ) |> 
  ggplot(aes(x = h, y = pct_change, color = .model)) +
    geom_line() +
    labs(x = "Horizon", y = "% improvements") +
    facet_wrap(~ level, ncol=2, scales = "free_y") +
    theme_bw() +
    scale_x_continuous(limits = c(1, 12), expand = c(0, 0)) +
    theme(legend.position = "bottom", legend.title = element_blank())
```

@@

visnights_full |> 
  filter(
    is_aggregated(Purpose) & is_aggregated(Zone) & is_aggregated(Region),
    !is_aggregated(State)
  ) |> 
  model(
    STL(Nights)
  ) |> 
  components() |> 
  pivot_longer(
    cols = c(Nights, trend, season_year, remainder, season_adjust),
    names_to = "component",
    values_to = "value"
  ) |>
  filter(
    component == c("Nights", "trend"),
    year(Month) >= 2000
  ) |>
  as_tibble() |>
  mutate(component = recode(component,                     # give them nicer names
                            Nights = "Observed Nights",
                            trend  = "STL Decomposed Trend")) |> 
  ggplot(aes(Month, value, color = as.character(State))) +
    geom_line() +
    facet_wrap(~ component, ncol=1, scales = "free_y") +
    labs(
      x = "Month",
      y = "Visitor Nights (in thousands)",
      color = "State"
    ) +
    theme_bw()
```

@@

# % improv plot by level
tourism_MSE_ex2016 |>
  group_by(series, h) |>
  mutate(
    base_MSE = MSE[.model == "base"]
  ) |>
  ungroup() |>
  group_by(.model, level, h) |>
  summarise(
    MSE = mean(MSE),
    base_MSE = mean(base_MSE),
    pct_change = (MSE - base_MSE) / base_MSE * 100
  ) |>
  filter(.model != "mint_sample") |>
  # Reorder the levels according to the desired order
  mutate(
    level = factor(
      level,
      levels = desired_order,
      labels = str_replace_all(desired_order, "_", " ")  # replace "_" with " "
    )
  ) |>
  ggplot(aes(x = h, y = pct_change, color = .model)) +
    geom_line() +
    labs(x = "Horizon", y = "% improvements") +
    facet_wrap(~ level, ncol=2, scales = "free_y") +
    theme_bw() +
    scale_x_continuous(limits = c(1, 12), expand = c(0, 0)) +
    theme(legend.position = "bottom", legend.title = element_blank())
```

@@

    "Final writing, presenation, submission"
  ),
  Deliverable = c(
    "Initial literature review + paper draft",
    "ReconCov package: Main codes for estimators + simulation framework + cross-validation",
    "NOVELIST initial simulation results + research proposal + presentation",
    "More simulation results on NOVELIST",
    "Empirical results on forecasting Tourism data",
    "New simulation design, POET & PC-adjusted NOVELIST simulation results",
    "Other estimators simulation results and full-fledged covariance estimators assessment",
    "Presentation and thesis manuscript"
  )
) |>
  kable(
    format = "latex",
    booktabs = TRUE,
    longtable = TRUE,
    caption = "Project Timeline and Key Milestones",
    escape = TRUE
  ) |>
  column_spec(1, width = "1.9cm") |>
  column_spec(2, width = "8cm") |>
  column_spec(3, width = "4.8cm") |>
  kable_styling(latex_options = "hold_position", font_size = 10)
```

@@

## Appendix: Australian Domestic Tourism Geographical Hierarchy {#sec-tourism-supplement}

```{r}
#| tbl-cap: "Geographical divisions of Australia."
#| label: tbl-tourism-geo
#| warning: false
#| message: false

source(paste0(path, "R/table_geo.R"))
table_geo_labeled <- create_table_geo()
table_geo_labeled[1,1] <- "Total"

df <- table_geo_labeled %>% 
  select(Series, Name, Label)

n  <- ceiling(nrow(df) / 2)
L  <- df[1:n, ]
R  <- df[(n+1):nrow(df), ]

# pad the shorter side
if (nrow(R) < nrow(L)) {
  R <- bind_rows(R, tibble(Series = rep(NA_integer_, n - nrow(R)),
                           Name = "", Label = ""))
}

wide <- bind_cols(
  setNames(L, c("Series","Name","Label")),
  setNames(R, c("Series","Name","Label"))
)

kable(
  wide,
  align     = c("r","l","l","r","l","l"),
  col.names = c("Series","Name","Label","Series","Name","Label")
) |> 
  kable_styling(latex_options = "hold_position", font_size = 8)
```
